Boutiques: a flexible framework for automated application integration in computing platforms

### 2016-10-20 paper.tex
More text

\author{Tristan Glatard, Tristan Aumentado-Armstrong, Natacha Beck, \\
  Pierre Bellec, Remi Bernard, Sorina Camarasu-Pop, \\
  Fr\'ed\'eric Cervenansky, Rafael Ferreira da Silva, \\ 
  John Flavin, Pascal Girard, Krzysztof J. Gorgolewski, \\
  Nathaniel Kofalt, Pierre-Olivier Quirion, Pierre Rioux, \\
  Gunnar Schaeffer, Marc-\'Etienne Rousseau, Alan C. Evans}

@@

Why Linux. Define application. Define computing platforms. Explain the
problem and why it is difficult. Approach: start simple and enrich the
spec when new applications come. Refer to computing infrastructures
and the need for parallelization. Define task vs. invocation. Who are
the developers. Refer to neuroinformatics field.

@@

\subsection{Workflow support}

@@

@@

\subsection{Core tools} Validator, local executor, invocation schema generator.

@@

\subsection{Ported pipelines}

% cbrain-plugins-neuro
% PSOM
% MICCAI Challenges
% exports from Nipype
% SPM?
% BIDS Apps

@@

@@

% CWL

>>>>

\author{Tristan Glatard, Tristan Aumentado-Armstrong, Natacha Beck, \\
  Pierre Bellec, Remi Bernard, Sorina Camarasu-Pop, \\
  Fr\'ed\'eric Cervenansky, Samir Das, Rafael Ferreira da Silva, \\ 
  John Flavin, Pascal Girard, Krzysztof J. Gorgolewski, \\
  Nathaniel Kofalt, Pierre-Olivier Quirion, Pierre Rioux, \\
  Gunnar Schaeffer, Marc-\'Etienne Rousseau, Alan C. Evans}

@@

Why Linux. Define application. Define computing platforms. Explain the
problem and why it is difficult. Approach: start simple and enrich the
spec when new applications come. Refer to computing infrastructures
and the need for parallelization. Who are
the developers. Refer to neuroinformatics field.

Vocabulary (check consistency across the paper):
\begin{itemize}
\item tool vs. application.
\item descriptor vs. manifest.
\item task vs. invocation.
\end{itemize}

@@

\subsection{Parallelization support}

@@

%\todo{rewrite this section to focus more on parallelism. Start by mentioning data parallelism.}

@@

\subsection{Core tools}

The Boutiques validator checks conformance of JSON manifests to the
Boutiques schema using a basic JSON validator. It also performs an
additional set of checks that cannot be easily implemented in JSON
schema: command-line keys are unique among inputs, input and output
identifiers are unique, input and output command-line keys are all
included in the command-line, command-line keys are not contained
within each other (which would puzzle substitution), output path
templates are unique (to avoid results to overwrite each other),
inputs of type Flag have a command-line flag, are optional and are not
lists, number and list constraints are sensible (e.g. min is lower
than max), the default value of enum types is part of the enum, an
input cannot both require and disable another input, required inputs
cannot require or disable other parameters, group member identifiers
must correspond to existing inputs and cannot appear in different
groups, mutually exclusive groups cannot have members requiring other
members, and that one-is-required groups should never have required
members. The Boutiques validator is meant for application developers.

The local executor can generate and execute command line from a
Boutiques manifest and a set of input values represented in a CSV or
JSON file complying with the invocation schema. It runs the
command in a container provided that the required framework
(e.g. Docker) is already installed. It can also generate hypothetical
command lines from random values, for debugging purposes. The local
executor can be used by application users, to run tools locally, or by
platforms, to generate command lines to be run on the execution
infrastructure.

The invocation schema generator creates an invocation schema (see
Section~\ref{sec:invocation-schema}) from a Boutiques manifest and
validates data against it. It can be used by platform administrators
to add invocation schemas to existing manifests, and by platforms to
validate input data. Validation is done using a basic JSON
validator.

@@

\subsection{Ported applications}

% cbrain-plugins-neuro

The Pipeline System for Octave and Matlab
(PSOM)~\cite{bellec2012pipeline} has been integrated in CBRAIN using
Boutiques. More precisely, a manifest has been written for a Niak
pre-processing pipeline and PSOM workers. The integration uses the
parallelization mechanism described in
Section~\ref{sec:parallelization} so that even the tasks processing a
given subject can be parallelized. As shown in~\cite{GLAT-16}, this
allows CBRAIN to leverage the efficient agent model used in PSOM. The
integration required some work in PSOM to facilitate its invocation as
a non-interactive command-line application. The resulting CBRAIN
plugin is available at
\url{https://github.com/SIMEXP/cbrain-plugins-psom}. It could in
principle be used in other platforms that support the parallelization
property, although this has not been tested.

% PSOM
% MICCAI Challenges 
% exports from Nipype
% SPM?
% BIDS Apps

@@

\begin{tabular}{ccccccccccccc}
  Application & \multicolumn{3}{c}{Containers} & \multicolumn{7}{c}{Input/Output properties}                         & \multicolumn{2}{c}{Other properties}  \\
              & Docker & Singularity & Rootfs  & Dependencies & Groups & Lists & Optional & Default & Enum & Min/max & Env vars & Min/Max \\

PSOM & 
\end{tabular}

@@

% CWL
The Common Workflow Language (CWL) is the closest related work to
Boutiques, in the sense that it also allows the description of command
lines executed in containers. According to Github, the CWL project
started 6 months before Boutiques. The main conceptual differences
between CWL and Boutiques are the following:
\begin{itemize}
\item CWL has a workflow language, Boutiques does not. CWL also has a
  command line tool description language, but its specification is
  influenced by the fact that command line tools are included in a
  specific workflow languages. In Boutiques, the only available
  workflow construct is sub-tasking (``a task may submit other
  tasks''). This is intentional as we envisage that Boutiques tools
  may be composed through any type of workflow engine and language. For instance, this leads to have multiple output types in CWL.
\item CWL specifies command lines formally whereas in Boutiques they
  are assembled from simple string templates. In short, a command line
  in CWL is specified by an executable and an ordered list of
  parameters. While this representation has the advantage of being
  more formal, for instance the executable can be identified
  automatically, it also removes flexibility by imposing that a tool
  only runs a single command. On the contrary, Boutiques allows any
  type of bash operators to be combined on the command line. This
  level of flexibility is important and greatly reduces the ``shim
  syndrome''. Boutiques' template approach allows for configuration files. 
\item CWL is validated using SALAD, which makes the ecosystem slightly
  more complex. For instance, CWL has ``import'' and ``include''
  directives. Boutiques is validated using plain JSON schema. \todo{ontology layer}.
\item In CWL, the applications need to be instrumented as follows, which is not the case in Boutiques:
\begin{itemize}
\item Applications must write in specific directories.
\item Applications must produce a JSON object describing the output structure.
\end{itemize}
\end{itemize}

\todo{Tristan}{Update from comparison summary on Google doc}
